A Wait-free Queue with Polylogarithmic Step Complexity

### 2023-01-05 podc_paper/algorithm-2.tex
Finished draft of analysis section.  (For some reason, compiling on my laptop requires commenting out some \newtheorem commands in podc-submission.tex .)
minor edits

\renewcommand{\algorithmiccomment}[1]{\hfill\eqparbox{COMMENTSINGLE}{\com\ #1}}
\begin{figure}
%\here{Does \var{leaf} have to be shared for helping when doing GC?}
\begin{algorithmic}[1]
\setcounter{ALG@line}{1}
\Statex \linecomment{Shared variable}
\begin{itemize}
\item \typ{Node} \var{root} \Comment{root of a binary tree of \tt{Node}s with one leaf for each process} % TODO: indent these lines in order to be more readable. I tries some times and i could not figure out doing that in algorithm.
\item \gc{\typ{*Node[]} \var{leafOfProcess[1..p]} \Comment{Pointer to the leaf assigned to each process.}}
\item \gc{\typ{int[]} \var{lastDequeuedByProcess[1..p]}}
\end{itemize}

@@

\renewcommand{\algorithmiccomment}[1]{\hfill\eqparbox{COMMENTDOUBLE}{\com\ #1}}
\newcommand{\hangbox}[1]{\parbox[t]{\dimexpr\linewidth-\algorithmicindent\relax}{%
    \setlength{\hangindent}{\algorithmicindent}%
    #1}}
\newcommand{\spac}{\medskip}
\newcommand{\assign}{\mbox{:=}}

@@

\gc{\If{ \var{v} is root and \var{h}\%p\textsuperscript{2}==0}
\State \Call{Help()}{}
\State \Call{FreeMemory}{root, MAX(lastDequeuedByProcess)-1}
\EndIf}

@@

\Function{element}{FindResponse}{\typ{int} \var{b}, \typ{int} \var{i}}
    \State \linecomment{find response to \var{i}th dequeue in \var{root}.\fld{blocks}[\var{b}]}
    \State \hangbox{\var{num\sub{enq}} \assign\ $\var{root}.\fld{blocks}[\var{b}].\fld{sum\sub{enq}} - \mbox{ }$\\
    		$\var{root}.\fld{blocks}[\var{b}-1].\fld{sum\sub{enq}}$}
    \If{$\var{root}.\fld{blocks}[\var{b}-1].\size + \var{num\sub{enq}} < \var{i}$}\label{checkEmpty}
    \gc{\State \var{lastDequeuedByProcess[this process id]} \assign\ \var{b}}
        \State \Return \nl \hfill \com\ queue is empty when dequeue occurs\label{returnNull}
            \EndIf
\State \linecomment{response is the \var{e}th enqueue in the root}
        \State \var{e} \assign\ \var{i} + \var{root}.\fld{blocks}[\var{b}-1].\fld{sum\sub{enq}} - 
			\var{root}.\fld{blocks}[\var{b}-1].\size\label{computeE}
		\State \linecomment{compute enqueue's block using binary search}
		\State find min $\var{b}'\leq \var{b}$ with $\var{root}.\fld{blocks}[\var{b}'].\fld{sum\sub{enq}} \geq \var{e}$
		\State \linecomment{find rank of enqueue within its block}
		\State $\var{i}' \assign\ \var{e} - \var{root}.\fld{blocks}[\var{b}'-1].\fld{sum\sub{enq}}$
        \State \var{e} \assign\ \Call{GetEnqueue}{\var{root}, $\var{b}'$, $\var{i}'$}\label{findAnswer}
        \gc{\State \var{lastDequeuedByProcess[this process id]} \assign\ $\var{b}'$}
        \State \Return{\var{e}}

@@

\begin{lemma}
    $\nf{Max(LastDequeuedFrom)} - \text{index of the last finished block}$ in the node $n$ is $O(p)$.
\end{lemma}

>>>>

\renewcommand{\algorithmiccomment}[1]{\hfill\eqparbox{COMMENTSINGLE}{\com\ #1}}
\begin{figure}
%\here{Does \var{leaf} have to be shared for helping when doing GC?}
\begin{algorithmic}[1]
\setcounter{ALG@line}{1}
\Statex \linecomment{Shared variable}
\begin{itemize}
\item \typ{Node} \var{root} \Comment{root of a binary tree of \tt{Node}s with one leaf for each process} % TODO: indent these lines in order to be more readable. I tries some times and i could not figure out doing that in algorithm.
\item \gc{\typ{*Node[]} \var{leafOfProcess}$[1..p]$ \Comment{Pointer to the leaf assigned to each process.}}
\item \gc{\typ{int[]} \var{last}$[1..p]$ \Comment{index of last block each process has dequeued an element from}}
\end{itemize}

@@

\renewcommand{\algorithmiccomment}[1]{\hfill\eqparbox{COMMENTDOUBLE}{\com\ #1}}

@@

\gc{    
	\If{ \var{v} is root and $\var{h} \mbox{ mod } p^2=0$}
	\State \var{max} \assign 0
	\For {$k \assign 1..p$}
		\State $\var{max} \assign \max(\var{max}, last[k])$
	\EndFor 
	\State \Call{Help()}{}
	\State \Call{FreeMemory}{$\var{root}, \var{max}-1$}
	\EndIf
}

@@

\Function{element}{FindResponse}{\typ{int} \var{b}, \typ{int} \var{i}}
    \State \linecomment{find response to \var{i}th dequeue in \var{root}.\fld{blocks}[\var{b}]}
    \State \hangbox{\var{num\sub{enq}} \assign\ $\var{root}.\fld{blocks}[\var{b}].\fld{sum\sub{enq}} - \mbox{ }$\\
    		$\var{root}.\fld{blocks}[\var{b}-1].\fld{sum\sub{enq}}$}
    \If{$\var{root}.\fld{blocks}[\var{b}-1].\size + \var{num\sub{enq}} < \var{i}$}\label{checkEmpty}
    \gc{\State \var{last[this process id]} \assign\ \var{b}}
        \State \Return \nl \hfill \com\ queue is empty when dequeue occurs\label{returnNull}
            \EndIf
\State \linecomment{response is the \var{e}th enqueue in the root}
        \State \var{e} \assign\ \var{i} + \var{root}.\fld{blocks}[\var{b}-1].\fld{sum\sub{enq}} - 
			\var{root}.\fld{blocks}[\var{b}-1].\size\label{computeE}
		\State \linecomment{compute enqueue's block using binary search}
		\State find min $\var{b}'\leq \var{b}$ with $\var{root}.\fld{blocks}[\var{b}'].\fld{sum\sub{enq}} \geq \var{e}$
		\State \linecomment{find rank of enqueue within its block}
		\State $\var{i}' \assign\ \var{e} - \var{root}.\fld{blocks}[\var{b}'-1].\fld{sum\sub{enq}}$
        \State \var{e} \assign\ \Call{GetEnqueue}{\var{root}, $\var{b}'$, $\var{i}'$}\label{findAnswer}
        \gc{\State \var{last[this process id]} \assign\ $\var{b}'$}
        \State \Return{\var{e}}

@@

\begin{lemma}
    $\nf{Max(Last)} - \text{index of the last finished block}$ in the node $n$ is $O(p)$.
\end{lemma}

### 2023-01-05 podc_paper/correctness.tex
minor edits

\here{need to update this outline later}
First, we define and prove some facts about blocks and the node's \nf{head} field. Then, we introduce  the linearization ordering formally. Next, we prove double \nf{Refresh} on a node is enough to propagate its children's new operations up to the node, which is used to prove (1). After this, we prove some claims about the size and operations of each block, which we use to prove the correctness of \nf{DoublingSearch()}, \nf{GetEnqueue()} and \nf{IndexDequeue()}. Finally, we prove the correctness of the way we compute the response of a dequeue, which establishes (2).

\subsection{Basic Properties}

@@

\subsection{Propagating Operations to the Root}
Next, we show two \op{Refresh}es suffice to propagate operations from a child to its parent.
We say that node $v$ \emph{contains} an operation if some block in $v.\fld{blocks}$ contains the operation.
\here{move this defn earlier and just recall it here?}
Once a block is added to a node, it remains there forever.  Thus, if $v$ contains an operation at some time, it contains the operation at all later times too.

@@

\subsection{Correctness of Helping Routines}

@@

\subsection{Linearizability}

>>>>

After proving some basic invariants in Section \ref{sec::basicProperties},
we show in Section \ref{sec::propagating} that performing a double refresh at each node
is sufficient to propagate an operation to the root.
In Section \ref{sec::helpingCorrect} we show that the \op{getEnqueue} and \op{indexDequeue}
routines correctly navigate through the tournament tree.
Finally, we show that the implementation is linearizable in Section \ref{sec::linearizability}.

\subsection{Basic Properties}
\label{sec::basicProperties}

@@

\subsection{Propagating Operations to the Root}
\label{sec::propagating}

Next, we show two \op{Refresh}es suffice to propagate operations from a child to its parent.
We say that node $v$ \emph{contains} an operation if some block in $v.\fld{blocks}$ contains the operation.
\here{move this defn earlier and just recall it here?}
Once a block is added to a node, it remains there forever.  Thus, if $v$ contains an operation at some time, it contains the operation at all later times too.

@@

\subsection{Correctness of Helping Routines}
\label{sec::helpingCorrect}

@@

\subsection{Linearizability}
\label{sec::linearizability}
